---
title: "The 2017 UK General Election"
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(dplyr)
```

## Load the data

```{r, warning = FALSE}
data <- read.csv("BritainElects.csv", skip = 1, header=T, stringsAsFactors = F)
data[,7:15] <- lapply(data[,7:15],function(x){as.numeric(gsub(",", "", x))})
data[is.na(data)] <- 0
con <- data %>% filter(X2017 == "Con")
lab <- data %>% filter(X2017 == "Lab")
nrow(con)
```

## Labour's weakest seats

Among seats won by Labour, how many might Con + UKIP have won?

```{r}
labmargin <- with(lab, Con + UKIP - Lab)
sum(labmargin > 0)
```

And how many would have been won by a margin of 1000?

```{r}
sum(labmargin > 1000)
```

Which seats are these?

```{r}
print(lab[which(labmargin > 1000), c(1,7,8,12)], row.names = FALSE)
```

## Tories weakest seats

```{r}
conmargin <- with(con, Lab + LDem + SNP - Con)
sum(conmargin > 0)
```

And how many would have been won by a margin of 1000?

```{r}
sum(conmargin > 1000)
```

Which seats are these?

```{r}
print(con[which(conmargin > 1000), c(1,7:10)], row.names = FALSE)
```

## Conclusion

It seems pretty safe to say the Conservatives would have **at least 20 fewer seats**, taking them well below the required majority, if Labour, Lib Dem, and SNP voters had coordinated.

**This is a failure of [First-past-the-post](https://en.wikipedia.org/wiki/First-past-the-post_voting) voting systems**.

It is very difficult to communicate with all voters supporting several different parties to get them to adopt a strategy.

**[Better voting systems](https://en.wikipedia.org/wiki/Instant-runoff_voting) accomplish the strategic outcome automatically** and should be a top priority for everyone who supports democracy.